---
title: "Data Vizualization Sample"
author: "Abdullah Al Mahmud"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---
\fontsize{12}{22}
\selectfont

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, echo=FALSE}
library(rvest) # for extracting data from web
library(tidyverse)
library(RColorBrewer)
library(ggmap)
library(maps)
library(mapdata)
library(gridExtra) # To show plots side by side
```

```{r, echo=FALSE}
world_coordinates <- map_data("world") 
wquake <- read_csv("../data/quakeworld.csv")
```

This document shows some extents made possible with the help of R programming in
data visualization. 


## Showing Location of Earthquakes around the World.

The dots represent the colors, while the colors represent the magnitude of the 
earthquakes; the redder an earthquake is, the greater it's magnitude. **The 
plot can also show regions colored by specific criteria. **


```{r, echo=FALSE, fig.height=4}
p1 <- ggplot() + 
  geom_map( 
    data = world_coordinates, map = world_coordinates, 
    aes(long, lat, map_id = region),
    fill = "grey"
  )+
  theme(aspect.ratio = 0.6)

p1 + 
  geom_point(data = wquake, 
             aes(x = longitude, 
                 y = latitude,
                # size = mag,
                 color = mag))+
  scale_color_gradient(high = "red", low = "yellow")+
  labs(title = "Earthquakes around the World in 2024 (Jan to Sep)",
       color = "Magnitude")+
  theme(axis.text.x=ggplot2::element_blank(),
        axis.ticks.x=ggplot2::element_blank(),
        axis.text.y=ggplot2::element_blank(),
        axis.ticks.y=ggplot2::element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(), # Remove grid lines
        panel.grid.minor = element_blank())

```

\newpage

## Producing Bar Diagram Using Web Data

Now we use data from the internet to make our plots. We get the data on 
most [goal scorers in soccer by country from Wikipedia](https://en.wikipedia.org/wiki/List_of_top_international_men%27s_football_goalscorers_by_country). The function `html_table()` from `rvest` makes the task
very effortless. 

```{r, echo=FALSE}
src <- "https://en.wikipedia.org/wiki/List_of_top_international_men%27s_football_goalscorers_by_country"
top_goals <- html_table(read_html(src), header = TRUE)[[1]]

top_goals$Player=factor(top_goals$Player, 
                        levels = top_goals$Player[order(top_goals$Goals)])

top_goals <- pivot_longer(data = top_goals[1:10, 1:5], cols = c(Goals, Caps))

top_goals %>% ggplot(aes(Player, value, fill = name))+
  geom_bar(stat = "identity", position = "dodge")+
  coord_flip()+
  scale_fill_brewer(palette = "Spectral", name = "Type")+
  labs(title = "Top Goal Scorers by Country", y = "No. of Goals/Caps", x = "Name")+
  geom_text(aes(label=paste(value, name)), 
            vjust=0.5, color="black", hjust = 1.1,
            position = position_dodge(0.9), size=3.5)
```

### Interpretation of the Chart

From the chart, it becomes evident that Cristiano Ronaldo scored the highest 
number of goals, totaling 135. The plot also shows he is the one having played
the maximum number of matches (217 caps). The adjacent bars reveal how the 
number of goals compare with number of caps. It becomes immediately clear that
Ference Puskas gets the credit of the highest scoring rate (84 goals in 85 
matches). As a consequence, we may in fact need to plot the scoring rate to
see the best performers. 

\newpage

## Showing Proportion of People on a Donut/Pie Chart

This is a nicer alternative to Pie Chart, both showing percentage or part of 
whole. 

```{r, echo=FALSE}
us_races <- tibble(
  Race = c("White", "Black", "Multiracial", "Asian", "Others"), 
  Percentage = c(61.6, 12.4, 10.2, 6, 9.8))
d1 <- us_races %>% ggplot(aes(2, Percentage, fill = Race))+
  geom_col(position = "stack", width = 1)+
  xlim(0.5, 2.5)+
  coord_polar("y")+
  theme_void()+
  geom_text(aes(label = paste0(Percentage, " % \n ", Race)), 
            position = position_stack(vjust = 0.5), size = 4)+
  scale_fill_brewer(palette = 10)+
  theme(legend.position="none")+
  labs(title = "Races among US Population")

pie1 <- us_races %>% ggplot(aes(x="", y = Percentage, fill = Race))+
  geom_bar(stat="identity")+
  coord_polar("y")+
  theme_void()+
  theme(legend.position = "none")+
  geom_text(aes(label = paste0(Percentage, " % \n ", Race)), 
            position = position_stack(vjust = 0.5), 
            size = 4, color = "black")+
  scale_fill_brewer(palette = 2)

grid.arrange(d1, pie1, ncol = 2)
```

### Interpretation of the Charts

The plots reveal that most people in USA belong to White race, Black people
come second among other. The Asians constitute least number of people. 

# Methods Used in The Document

The charts in this documents and the document itself are generated within `R` 
programming language with the help of `Rmarkdown`. You can see the source code 
of this document along with the codes of the chart [here (CLICK to see)](https://github.com/mahmudstat/open-analysis/blob/main/sample/viz.Rmd). 

An updated (if any) version of this file available [here](https://github.com/mahmudstat/open-analysis/blob/main/sample/viz.pdf)

