---
title: "Data Vizualization Sample"
author: "Mahmud"
date: "2022-06-06"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, echo=FALSE}
library(rvest) # for extracting data from web
library(tidyverse)
library(RColorBrewer)
library(ggmap)
library(maps)
library(mapdata)
```

```{r, echo=FALSE}
world_coordinates <- map_data("world") 
wquake <- read_csv("../data/quakeworld.csv")
```

This document shows some extents made possible with the help of R programming in
data visualization. 

# Methods

The charts in this page are generated within `R` programming language with the
help of `Rmarkdown`. You can see the source code of this document along with 
the codes of the chart [here (CLICK to see)]. 

## Showing Location of Earthquakes around the World.

The dots represent the colors, while the colors represent the magnitude of the 
earthquakes; the redder an earthquake is, the greater it's magnitude. 

```{r, echo=FALSE}
p1 <- ggplot() + 
  geom_map( 
    data = world_coordinates, map = world_coordinates, 
    aes(long, lat, map_id = region),
    fill = "grey"
  )+
  theme(aspect.ratio = 0.6)

p1 + 
  geom_point(data = wquake, 
             aes(x = longitude, 
                 y = latitude,
                # size = mag,
                 color = mag))+
  scale_color_gradient(high = "red", low = "yellow")+
  labs(title = "Earthquakes around the World in 2024 (Jan to Sep)",
       color = "Magnitude")+
  theme(axis.text.x=ggplot2::element_blank(),
        axis.ticks.x=ggplot2::element_blank(),
        axis.text.y=ggplot2::element_blank(),
        axis.ticks.y=ggplot2::element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(), # Remove grid lines
        panel.grid.minor = element_blank())
```

\newpage

## Showing Proportion of People on a Donut Chart

This is a nicer alternative to Pie Chart, both showing percentage or part of 
whole. 

```{r, echo=FALSE}
us_races <- tibble(
  Race = c("White", "Black", "Multiracial", "Asian", "Others"), 
  Percentage = c(61.6, 12.4, 10.2, 6, 9.8))
us_races %>% ggplot(aes(2, Percentage, fill = Race))+
  geom_col(position = "stack", width = 1)+
  xlim(0.5, 2.5)+
  coord_polar("y")+
  theme_void()+
  geom_text(aes(label = paste0(Percentage, " % \n ", Race)), 
            position = position_stack(vjust = 0.5), size = 4)+
  scale_fill_brewer(palette = 10)+
  theme(legend.position="none")+
  labs(title = "Races among US Population")
```

The plot reveals that most people in USA belong to White race, Black people
come second among other. The Asians constitute least number of people. 

\newpage

## Producing Bar Diagram Using Web Data

Now we use data from the internet to make our plots. We get the data on 
most [goal scorers in soccer by country from Wikipedia](https://en.wikipedia.org/wiki/List_of_top_international_men%27s_football_goalscorers_by_country). The function `html_table()` from `rvest` makes the task
very effortless. 

```{r, echo=FALSE}
src <- "https://en.wikipedia.org/wiki/List_of_top_international_men%27s_football_goalscorers_by_country"
top_goals <- html_table(read_html(src), header = TRUE)[[1]]

top_goals$Player=factor(top_goals$Player, 
                        levels = top_goals$Player[order(top_goals$Goals)])

top_goals <- pivot_longer(data = top_goals[1:10, 1:5], cols = c(Goals, Caps))

top_goals %>% ggplot(aes(Player, value, fill = name))+
  geom_bar(stat = "identity", position = "dodge")+
  coord_flip()+
  scale_fill_brewer(palette = "Spectral", name = "Type")+
  labs(title = "Top Goal Scorers by Country", y = "No. of Goals/Caps", x = "Name")+
  geom_text(aes(label=paste(value, name)), 
            vjust=0.5, color="black", hjust = 1.1,
            position = position_dodge(0.9), size=3.5)
```

From the chart, it becomes, 